Differentiable Optics with {\Large $\partial$}Lux: I - Deep Calibration of Flat Field and Phase Retrieval with Automatic Differentiation

### 2022-12-16 src/tex/ms.tex
[showyourwork] overleaf sync

The wavefront then encounters low-order phase errors parameterised by coefficients on Zernike polynomials. We include the first 4 radial terms, ignoring global piston, tip and tilt for a total of 7 terms. Coefficients were randomly drawn from a normal distribution of mean 0 and standard deviation 20\,nm, for a total of 
\textcolor{red}{showyourwork variable}
% \variable{output/rms_opd_in.txt}\,nm 
RMS total \ac{opd}. These aberrations applied to the pupil mask and resulting \ac{psf} are shown in the middle panel of Figure~\ref{fig:optics}, where we see this induces a large amount of distortion in the \ac{psf}.

@@

% Describe gradient descent implementation & scheduling
% Takes X time on Y machine with Z precision on params (numbers from ShowYourWork)
In order to recover these input aberrations, we perform gradient descent using Adam \cite{Kingma2014} as implemented in \optax \cite{optax2020github}, minimizing the posterior of a per-pixel Poisson log-likelihood and a $\chi^2$ log prior on the PRF values, matching the true distribution of mean 1 deviation 0.05. We take the gradient of this loss function with respect to the position and flux of all stars, the response in all pixels, and the Zernike coefficients. We also use a staged optimisation strategy, initially optimising the positions, fluxes and optical aberrations in order to recover a good \ac{psf} model before learning the \ac{prf}. This is necessary to circumvent the high degree of covariance that each individual pixel experiences with the parameters of the \ac{psf}. To achieve this the learning rate of the \ac{prf} is set to zero for the first 50 epochs of optimisation.

@@

This optimisation strategy performs well, with all of the parameters being recovered down to the noise level. This can be seen in Figure~\ref{fig:data_residual} with the right panel showing a sample residual of the data and model after optimisation, which is clearly only noise. Correlations and residual plots of the positions and fluxes after the optimisation are shown in Figure~\ref{fig:astro_params}. The 1-$sigma$ error bars indicate a poor recovery, however these are only marginalised over positions, fluxes and aberration, not the \ac{prf} which is the dominant noise source. These parameters are recovered to the expected distribution when performed without modelling the \ac{prf}. Figure~\ref{fig:aberrations} shows the recovery of the optical aberrations, with a final residual aberration of 
\textcolor{red}{showyourwork variable}
% \variable{output/rms_opd_resid.txt} nm RMS. 
Finally, Figure~\ref{fig:flat_field} shows a correlation plot of the true and recovered sensitivity values and the corresponding residual histogram. The points of the correlation plot are colour-coded with the total counts for each pixel, showing as expected that pixels with higher photon counts having a better recovery of the PRF since these pixels have a greater \ac{snr}. Overall the parameters are very well recovered down to the noise level.

>>>>

The wavefront then encounters low-order phase errors parameterised by coefficients on Zernike polynomials. We include the first 4 radial terms, ignoring global piston, tip and tilt for a total of 7 terms. Coefficients were randomly drawn from a normal distribution of mean 0 and standard deviation 20\,nm, for a total of 
% \textcolor{red}{showyourwork variable}
\variable{output/rms_opd_in.txt}\,nm 
RMS total \ac{opd}. These aberrations applied to the pupil mask and resulting \ac{psf} are shown in the middle panel of Figure~\ref{fig:optics}, where we see this induces a large amount of distortion in the \ac{psf}.

@@

% Describe gradient descent implementation & scheduling
% Takes X time on Y machine with Z precision on params (numbers from ShowYourWork)
In order to recover these input aberrations, we perform gradient descent using Adam \cite{Kingma2014} as implemented in \optax \cite{optax2020github}, minimizing the posterior of a per-pixel Poisson log-likelihood and a $\chi^2$ log prior on the PRF values, matching the true distribution of mean 1 deviation 0.05. We take the gradient of this loss function with respect to the position and flux of all stars, the response in all pixels, and the Zernike coefficients. We also use a staged optimisation strategy, initially optimising the positions, fluxes and optical aberrations in order to recover a good \ac{psf} model before learning the \ac{prf}. This is necessary to circumvent the high degree of covariance that each individual pixel experiences with the parameters of the \ac{psf}. To achieve this the learning rate of the \ac{prf} is set to zero for the first 50 epochs of optimisation, after which the \ac{prf} learning rate is initialised while the position, flux and aberrations parameters are frozen for 25 epochs before being allowed to vary again.

@@

This optimisation strategy performs well, with all of the parameters being recovered down to the noise level. This can be seen in Figure~\ref{fig:data_residual} with the right panel showing a sample residual of the data and model after optimisation, which is clearly only noise. Correlations and residual plots of the positions and fluxes after the optimisation are shown in Figure~\ref{fig:astro_params}. The 1-$sigma$ error bars indicate a poor recovery, however these are only marginalised over positions, fluxes and aberration, not the \ac{prf} which is the dominant noise source. These parameters are recovered to the expected distribution when performed without modelling the \ac{prf}. Figure~\ref{fig:aberrations} shows the recovery of the optical aberrations, with a final residual aberration of 
% \textcolor{red}{showyourwork variable}
\variable{output/rms_opd_resid.txt} nm RMS. 
Finally, Figure~\ref{fig:flat_field} shows a correlation plot of the true and recovered sensitivity values and the corresponding residual histogram. The points of the correlation plot are colour-coded with the total counts for each pixel, showing as expected that pixels with higher photon counts having a better recovery of the PRF since these pixels have a greater \ac{snr}. Overall the parameters are very well recovered down to the noise level.
